---
title: "MNCWTA Text Mining"
author: "Patrick Torralba"
format: 
  html:
    theme: flatly
    fontsize: 0.8em
    code-fold: FALSE
    code-tools: TRUE
    code-copy: TRUE
    toc: TRUE
    toc-depth: 4
    smooth-scroll: TRUE
    embed-resources: TRUE
    fig-align: center
editor: visual
---

# Library

```{r}
#| label: library
#| message: false
#| warning: false
#| echo: true  

library(pdftools) # to extract data from practice framework .pdf file
library(tidyverse) # pull strings
```

# Import MN Practice Framework

```{r}
#| label: import
#| message: false
#| warning: false
#| echo: true  

pdf_raw <- pdf_text("MN Child Welfare Framework for Competent Practice.pdf")

```

# Data Cleaning

This is the prompt I used in gemini to pull data:

"how to extract text from tables in a .pdf document in R?"

## Page Extraction and table conversion

*Explanation*

-   We first specify the pages we want
-   then we tell R to split the strings based on the line break "\n"
-   unlist() takes the list and gives us 1 singular column
-   then we convert to a table
-   and we also perform some minor data cleaning like remove the excess white space and blank rows

```{r}
#| label: page-extraction-cleaning
#| message: false
#| warning: false
#| echo: true  

# get competencies only
comp_table <- pdf_raw[c(8:25)] %>% 
  str_split("\n") %>% 
  unlist() %>% 
  as_tibble() %>% 
  slice(-c(1:4)) %>% ## start at the 5th row since the first 4 rows are just fluff
  mutate(value = str_trim(value)) %>% 
  filter(!is.na(value),
         value != "")


```

-   clean the raw text data by removing side paragraph

This is essentially saying create a new column called "text" with the following logic: "\\s{2,}" if a row contans 2+ spaces then remove everything left up to the gap

And we also trim the column at the end

```{r}
#| label: page-table-cleaning
#| message: false
#| warning: false
#| echo: true  

comp_table_2 <- comp_table %>% 
  mutate(text = if_else(
    condition = str_detect(value, "\\s{2,}"),
    true = str_remove(value, "^.*?(?=\\s{2,})"),
    false = value)) %>% 
  mutate(text = str_trim(text))
    

```

-   Now in this step, we take bind rows together. Namely, we bind the current and next row together until it hits a digit value.

-   is_header is a column that checks if the row starts with a digit followed by a letter and a dot (e.g "1a.")

-   then we create a unique group_id for each learning objective to track the text

-   afterwards, we combine the columns using the summarize() and paste() functions

-   lastly, we remove irrelevant words from the learning objectives. We do this by removing any character after the ")" character.

```{r}
#| label: learning-objectives cleaning
#| message: false
#| warning: false
#| echo: true  

comp_table_3 <- comp_table_2 %>% 
  select(text) %>% 
  mutate(is_header = str_detect(text, "^\\d+[a-z]\\.")) %>% 
  # Create a unique group ID that increments every time 'is_header' is TRUE
  mutate(group_id = cumsum(is_header)) %>%
  # Group by this new ID and collapse the text rows together
  group_by(group_id) %>%
  summarise(
    combined_text = paste(text, collapse = " "),
    .groups = "drop"
  ) %>%
  # Clean up by removing the temporary group_id column if you don't need it
  select(-group_id) %>% 
  mutate(clean_text = str_remove(combined_text, "(?<=\\)).*$")) %>% 
  select(clean_text)
  
  
```

# Text Mining Notes

-   str_detect(text, "\^\\d+\[a-z\]\\."): This looks at the beginning of the string (\^). It checks for one or more digits (\\d+), followed by a lowercase letter (\[a-z\]), followed by a literal period (\\.).
-   cumsum(is_header): In R, TRUE equals 1 and FALSE equals 0. By taking the cumulative sum of the TRUE values, R creates a grouping variable that stays the same until it hits the next TRUE, where it bumps up by 1.
-   paste(text, collapse = " "): Once grouped by that ID, paste(..., collapse = " ") stitches all the character strings in that group into a single string, separated by a space.
-   Look-up regex cheat sheet to learn how to read this information.

## Regex Tips

## 📝 Text Mining & Regex Cheat Sheet (R / `stringr`)

### 1. Core Lookbehind Concepts

A **lookbehind assertion** is a zero-width match that checks the text to the left *before* allowing a match, without including that text in the final result.

> ⚠️ **The Golden Rule:** Lookbehinds in R (`stringr`/`stringi`) **must have a fixed length**. You cannot use variable quantifiers like `*` or `+` inside them.

| Type                    | Syntax         | Meaning                                | R Example Pattern                                  |
|:-----------------|:-----------------|:-----------------|:-----------------|
| **Positive Lookbehind** | `(?<=pattern)` | Match if `pattern` is on the left.     | `(?<=\\$)\\d+` (Extracts digits preceded by `$`)   |
| **Negative Lookbehind** | `(?<!pattern)` | Match if `pattern` is NOT on the left. | `(?<!-)\\d+` (Extracts digits NOT preceded by `-`) |

------------------------------------------------------------------------

### 2. Common `stringr` Functions

Always stick to `stringr` functions for consistent `(string, pattern)` syntax.

-   `str_detect(string, pattern)`: Returns `TRUE` / `FALSE`.
-   `str_extract(string, pattern)`: Pulls out the *first* match.
-   `str_remove(string, pattern)`: Deletes the matched pattern.
-   `str_replace(string, pattern, replacement)`: Replaces the match.

------------------------------------------------------------------------

### 3. Regex Syntax Building Blocks

#### The Double Backslash (`\\`)

In R, you must use a **double backslash** to escape special regex characters when you want to match them literally. \* Literal Period: `\\.` \* Literal Parentheses: `\\(` or `\\)` \* Literal Question Mark: `\\?`

#### Anchors & Classes

| Regex Component | What it Matches                           | Example                                   |
|:-----------------------|:-----------------------|:-----------------------|
| `^`             | Start of the string                       | `^\\d` (Starts with a digit)              |
| `$`             | End of the string                         | `\\.$` (Ends with a period)               |
| `\\d`           | Any digit (0-9)                           | `\\d{4}` (Exactly 4 digits)               |
| `\\w`           | Any word character (letters/numbers)      | `\\w+` (One or more characters)           |
| `\\s`           | Any whitespace (space, tab, newline)      | `\\s+` (One or more spaces)               |
| `.*`            | Anything (Any character, 0 or more times) | `\\).*$` (Everything after a parenthesis) |

------------------------------------------------------------------------
