---
title: "SFI Table Text Mining"
author: "Patrick Torralba"
format: 
  html:
    theme: flatly
    fontsize: 0.8em
    code-fold: FALSE
    code-tools: TRUE
    code-copy: TRUE
    toc: TRUE
    toc-depth: 4
    smooth-scroll: TRUE
    embed-resources: TRUE
    fig-align: center
editor: visual
---

# Library
```{r}
#| label: library
#| message: FALSE
#| warning: FALSE
#| echo: TRUE 

library(tidyverse) #clean data
library(pdftools) # manipulate strings
```

# import 
```{r}
#| label: import
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

SFI_raw <- pdf_text("Operationalizing Kadushin s Model of Social Work Supervision.pdf")
```

# Data cleaning

## Page Extraction
I only want to extract pages 8-10 because they contain the SFI Action Items
```{r}
#| label: page-extraction
#| message: FALSE
#| warning: FALSE
#| echo: TRUE    

SFI <- SFI_raw[c(8:10)] |> 
  str_split("\n") |> 
  unlist() |> 
  as_tibble() |> 
  slice(-c(1:11)) |>  #start at the 11th row bc the rest is j fluff
  mutate(value = str_trim(value)) |> 
  filter(!is.na(value),
         value != "") |> 
  slice(-c(60:n())) #drop rows after row 59 bc they are outside of our SFI table
```

## Page Cleaning

This is my logic in cleaning the pages to effectively extract SFI information:

determine if a row contains a capital letter [this tells us the start and end of a sentence] --> If Yes = group ID, no = no group ID and keep searching until the next iteration
* This is so I can determine where the sentence starts and ends since the data got split by row-level

I do a bit more data cleaning and create a variable where I check if the text variable starts with a capital letter or not. I am doing this so I can create a unique ID.
```{r}
#| label: SFI-cleaning
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

SFI_2 <- SFI |> 
  mutate(value = str_squish(value)) |> 
  filter(!str_detect(value, "Irit"),
         value != "TABLE 1 (continued)",
         value != "116 JOURNAL OF SOCIAL SERVICE RESEARCH",
         value != "116             JOURNAL OF SOCIAL SERVICE RESEARCH")|> 
  mutate(capital_letter = str_detect(value, "^[A-Z]")) # this checks if the row starts with a capital letter
```

Now that I've determined which rows contain capital letters I'm going to create a group ID and merge rows together to make coherent SFI action items.

```{r}
#| label: SFI-group-id
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

SFI_3 <- SFI_2 |> 
  mutate(group_id = cumsum(capital_letter)) |> 
  group_by(group_id) |> 
  summarize(combined_text = paste(value, collapse = " "))

```

I'm going to export a .csv so I can review what else I need to clean!

```{r}
#| label: SFI-3-export
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

write.csv(SFI_3, "SFI_3.csv")
```


Now I'm j going to continue data cleaning!

```{r}
#| label: SFI-cleaning-giant-case-when
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

SFI_4 <- SFI_3 |> 
  filter(combined_text != "Factors Action Items Loadings") |> 
  mutate(combined_text_2 = case_when( # this removes extraneous text from the Factors side paragraph
    combined_text ==  "Policy, Planning, Discussion with workers about plans & Budgeting implemented in the community or in the agency" ~ "Discussion with workers about plans implemented in the community or in the agency",
    combined_text == "& Budgeting implemented in the community or in the agency" ~ "implemented in the community or in the agency",
    combined_text == "Controllingthat assignment are completed on time" ~ "Controlling that assignment are completed on time",
    combined_text == "Quality Control Regulating workers' case loads" ~ "Regulating workers' case loads",
    combined_text == "Contacts Finding or recruiting volunteers with Community Referrals to, and coordination or supervision" ~ "Finding or recruiting volunteers with Community Referrals to, and coordination or supervision",
    combined_text == "Services of private-service providers" ~ "of private-service providers",
    combined_text == "Professional Discussing assessment of clients with workers" ~ "Discussing assessment of clients with workers",
    combined_text == "Skills & Discussing intervention plans with workers" ~ "Discussing intervention plans with workers",
    combined_text == "Techniques Discussing intervention methods with workers (i.e., casework, group-work)" ~ "Discussing intervention methods with workers (i.e., casework, group-work)",
    combined_text == "Professional Discussing priorities regarding receiving clients" ~ "Discussing priorities regarding receiving clients",
    combined_text == "Knowledge and Discussion with workers about their working" ~ "Discussion with workers about their working",
    combined_text == "Information relations with interdisciplinary teams" ~ "relations with interdisciplinary teams",
    combined_text == "Referringworkers to information resources in the field/wmmunity (e.g., experts, reports)" ~ "Referring workers to information resources in the field/wmmunity (e.g., experts, reports)",
    combined_text == "Assisting workers in handling tensions and" ~ "Assisting workers in handling tensions and frustrations relating to their work in general",
    combined_text == "Finding or recruiting volunteers with Community Referrals to, and coordination or supervision" ~ "Finding or recruiting volunteers",
    combined_text == "of private-service providers" ~ "Referrals to, and coordination or supervision",
      .default = combined_text)) |> 
  select(combined_text_2)
  
  

```

I'm going to rinse and repeat the group_id process so I have complete coherent SI action items

```{r}
#| label: SFI-group-id-2
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

SFI_5 <- SFI_4 |> 
  mutate(capital_letter = str_detect(combined_text_2, "^[A-Z]")) |> 
  group_by(group_id = cumsum(capital_letter)) |> 
  summarize(SFI_clean = paste(combined_text_2, collapse = " "))


```

Almost done. Now I'm going to clean-up the remaining dirty rows by removing any digits that were accidentally carried over

```{r}
#| label: SFI-number-remove
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

SFI_6 <- SFI_5 |> 
  mutate(SFI_clean_2 = str_remove(SFI_clean, "\\.\\d+")) |> ## this removes the decimal point and all digits following it
  select(SFI_clean_2)

```

# Text Mining Notes
* \\. finds the literal decimal point.

* ".*" finds absolutely any character (.) that comes after it, any number of times (*), effectively clearing out the rest of the string.

# Export
Done! Now I'm going to export the finished product as a .csv to work on it 

```{r}
#| label: SFI-export-final
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

write.csv(SFI_6, "SFI_clean.csv")

```

