---
title: "New Supervisor Learning Objectives Merge"
format: 
  html:
    theme: flatly
    fontsize: 0.8em
    code-fold: FALSE
    code-tools: TRUE
    code-copy: TRUE
    toc: TRUE
    toc-depth: 4
    smooth-scroll: TRUE
    embed-resources: TRUE
    fig-align: center
editor: visual
---

Note: This is the improved version that contains the missing learning objectives in the first iteration of this .qmd file.

# library

```{r}
#| label: library
#| message: false
#| warning: false
#| echo: true  

library(tidyverse)
library(readxl)
```

# import data

```{r}
#| label: import
#| message: false
#| warning: false
#| echo: true  

learn_obj <- read_excel("New Supervisor Learning Objectives.xlsx")
comp_2 <- read_csv("Competency_Table_v2.csv")

```

# Data Cleaning

## Competency

-   We only want to extract New Supervisor information

```{r}
#| label: clean-competency
#| message: false
#| warning: false
#| echo: true  

comp_2 <- comp_2 %>% 
  filter(!is.na(`Supervisor Competency`)) %>% 
  select(`Competency Code`, `Competency Description`, Competencies, `Knowledge / Skill Statement`)
```

## Competency Code

-   In this step, we split the competency codes
    -   This is necessary for merging on a later step

```{r}
#| label: learn-obj-split
#| message: false
#| warning: false
#| echo: true  

learn_obj_2 <- learn_obj %>% 
  mutate(comp_code_org = learn_obj$`Competency Code`) %>% 
  separate_wider_delim(
    cols = `Competency Code`,
    delim = ",",
    names_sep = "Column",
    too_few = "align_start"
  )

```

### Merge

-   I merge the competency codes to the competency codes present in the learning objectives

```{r}
#| label: learn-obj-merge
#| message: false
#| warning: false
#| echo: true  

library(dplyr)
library(tidyr)


learn_obj_3 <- learn_obj_2 %>%
  # 1. Temporarily pivot columns 1 to 13 into a single long column
  pivot_longer(
    cols = starts_with("Competency CodeColumn"),
    names_to = "Original_Column",
    values_to = "Competency Code"
  ) %>%
  mutate("Competency Code" = str_trim(`Competency Code`, "both")) %>% 
  # 2. Do the single, clean join
  left_join(
    comp_2, 
    by = "Competency Code"
  ) %>%
  # 3. Pivot it back to the original wide format
  pivot_wider(
    names_from = Original_Column,
    values_from = c(`Competency Code`, `Competency Description`),
    names_glue = "{.value}_{Original_Column}"
  ) %>% 
  select(-starts_with("Competency Code_"))
  

```


#### Merge Cleaning

-   There are duplicate learning objectives due to varying competencies
  - Thus, I set up a group_id column and then bind similar rows
  
```{r}
#| label: bind-competency-row
#| message: false
#| warning: false
#| echo: true  


```

