---
title: "New Supervisor Learning Objectives Merge"
format: 
  html:
    theme: flatly
    fontsize: 0.8em
    code-fold: FALSE
    code-tools: TRUE
    code-copy: TRUE
    toc: TRUE
    toc-depth: 4
    smooth-scroll: TRUE
    embed-resources: TRUE
    fig-align: center
editor: visual
---

# library

```{r}
#| label: library
#| message: false
#| warning: false
#| echo: true  

library(tidyverse)
library(readxl)
```

# import data

```{r}
#| label: import
#| message: false
#| warning: false
#| echo: true  

learn_obj <- read_excel("New Supervisor Learning Objectives.xlsx")
comp <- read_csv("Competency_Table.csv")

```

# Data Cleaning

## Competency

-   We only want to extract New Supervisor information

```{r}
#| label: clean-competency
#| message: false
#| warning: false
#| echo: true  

comp <- comp %>% 
  filter(!is.na(`Supervisor Competency`))
```

## Competency Code

-   In this step, we split the competency codes
    -   This is necessary for merging on a later step

```{r}
#| label: learn-obj-split
#| message: false
#| warning: false
#| echo: true  

learn_obj_2 <- learn_obj %>% 
  mutate(comp_code_org = learn_obj$`Competency Code`) %>% 
  separate_wider_delim(
    cols = `Competency Code`,
    delim = ",",
    names_sep = "Column",
    too_few = "align_start"
  )

```

### Merge

-   I merge the competency codes to the competency codes present in the learning objectives

```{r}
#| label: learn-obj-merge
#| message: false
#| warning: false
#| echo: true  

library(dplyr)
library(tidyr)

##NEED TO FIX

learn_obj_3 <- learn_obj_2 %>%
  # 1. Temporarily pivot columns 1 to 13 into a single long column
  pivot_longer(
    cols = starts_with("Competency CodeColumn"),
    names_to = "Original_Column",
    values_to = "Competency Code"
  ) %>%
  mutate("Competency Code" = str_trim(`Competency Code`, "both")) %>% 
  # 2. Do the single, clean join
  left_join(
    comp %>% select(`Competency Code`, `Competency Description`), 
    by = "Competency Code"
  ) %>%
  # 3. Pivot it back to the original wide format
  pivot_wider(
    names_from = Original_Column,
    values_from = c(`Competency Code`, `Competency Description`),
    names_glue = "{.value}_{Original_Column}"
  ) %>% 
  select(-starts_with("Competency Code_"))
  

```

#### Locate Missing Data

We locate the missing learning objective information by filtering competency description column to NA or empty. \* I then add these missing learning objectives to the original comp \* I intend to re-merge again to get the missing LOs

```{r}
#| label: learn-obj-missing
#| message: false
#| warning: false
#| echo: true  

learn_obj_missing <- learn_obj_3 %>% 
  filter(is.na(`Competency Description_Competency CodeColumn1`) & !is.na(comp_code_org))# %>% 


```

#### Add Missing LO Descriptions to comp

I define the missing learning objectives and add them to the data frame \* Note: check.names = FALSE forces R to keep the exact column names intact

```{r}
#| label: comp-def-manipulate
#| message: false
#| warning: false
#| echo: true  

## make new df and define missing LO

missing_def <- data.frame("Competency Code" = c("IDC","IDP","4c-d","4k-l","1aa-bb","2k-l","1m-n"),
                          "Competency Description" = c("Intercultural Development Continuum", #IDC
                                      "Intercultural Development Plan", #IDP
                                      "Works with others to find a win/win resolution of differences whenever possible. (S); Shows awareness of how management style impacts staff productivity and development. (K)", #4c-d
                                      "Recognizes the complexities inherent in conflict, including aspects of direct child welfare practice, and proactively engages others in conflict resolution. (K&S); Understands the concepts of team development, facilitation of effective meetings, and conflict management. (K)", #4k-l
                                      "Plans, schedules, and directs the work of self and others. (S); Demonstrates ability to structure supervisory conferences (individual and group) to review and document casework activities and caseworker performance. (S)", #1aa-bb
                                      "Shows the ability to manage the employee appraisal process, to use competencies in performance planning, and to assess employee performance. (K&S); Demonstrates the ability to identify employee performance problems and apply principles of progressive discipline. (S)", #2k-l
                                      "Understands the value of supporting expression, and can recognize emotion-laden issues or situations and handle them with sensitivity. (S); Uses listening and observation to anticipate and prepare for others’ reactions. (K&S)"), #1m-n
                          "Competencies" = c(
                            NA, #IDC
                            NA, #IDP
                            "Team Leadership", #4c-d
                            "Team Leadership", #4k-l
                            "Administrative Supervision", #1aa-bb
                            "Human Resource Management", #2k-l
                            "Supportive Supervision" #1m-n
                          ),
                          "Supervisor Competency" = c(
                            NA, #IDC
                            NA, #IDP
                            "Team Leadership", #4c-d
                            "Team Leadership", #4k-l
                            "Administrative Supervision", #1aa-bb
                            "Human Resource Management", #2k-l
                            "Supportive Supervision" #1m-n
                          ),
                          "Knowledge / Skill Statement" = c(
                            NA, #IDC
                            NA, #IDP
                            "Skills; Knowledge", #4c-d
                            "Knowledge and Skills; Knowledge", #4k-l
                            "Skills; Skills", #1aa-bb
                            "Knowledge and Skills; Skills", #2k-l
                            "Skills; Knowledge and Skills"), #1m-n
                          check.names = FALSE)



```

Now I add the missing learning objective definitions to the data frame

```{r}
#| label: add-missing-def-to-comp
#| message: false
#| warning: false
#| echo: true  

comp_2 <- bind_rows(comp, missing_def)

```

I export this updated table incase I need it for the future.

```{r}
#| label: export-comp_v2
#| message: false
#| warning: false
#| echo: true  


write_csv(comp_2, "Competency_Table_v2.csv")
```

### Merge Missing Data back to original
