---
title: "New Supervisor Learning Objectives - Facilitator Guide"
format: 
  html:
    theme: flatly
    fontsize: 0.8em
    code-fold: FALSE
    code-tools: TRUE
    code-copy: TRUE
    toc: TRUE
    toc-depth: 4
    smooth-scroll: TRUE
    embed-resources: TRUE
    fig-align: center
editor: visual
---

Note: This version analyzes new supervisor learning objectives located in the Curriculum Library - Facilitation Guides.

# library

```{r}
#| label: library
#| message: false
#| warning: false
#| echo: true  

library(tidyverse)
```

# import data

```{r}
#| label: import
#| message: false
#| warning: false
#| echo: true  

learn_obj <- read_csv("Imports/New Supervisor Learning Objectives_v2 - Original_hide.csv") |> 
  select(`Learning Objective`)

comp_2 <- read_csv("Exports/Competency_Table_v2.csv")

```

# Data Cleaning

## Competency

- We only want to extract New Supervisor information

```{r}
#| label: clean-competency
#| message: false
#| warning: false
#| echo: true  

comp_2 <- comp_2 |> 
  filter(!is.na(`Supervisor Competency`)) |> 
  select(`Competency Code`, `Competency Description`, Competencies, `Knowledge / Skill Statement`)
```

## Competency Code

- In this step, we split the competency codes from the learning objectives
- This is necessary for merging on a later step

```{r}
#| label: learn-obj-comp-code
#| message: false
#| warning: false
#| echo: true  

learn_obj_2 <- learn_obj |> 
  mutate(comp_code = str_extract(`Learning Objective`, "\\(.*"))

```

```{r}
#| label: learn-obj-clean
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

learn_obj_2 <- learn_obj_2 |> 
  mutate(comp_code = str_remove(comp_code, "\\("),
         comp_code = str_remove(comp_code, "\\)"))


```

```{r}
#| label: learn-obj-split
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

learn_obj_2 <- learn_obj_2 |> 
  separate_wider_delim(
    cols = comp_code,
    delim = ",",
    names_sep = "_",
    too_few = "align_start")

```

### Merge

- I merge the competency codes to the competency codes present in the learning objectives

```{r}
#| label: learn-obj-merge
#| message: false
#| warning: false
#| echo: true  

library(dplyr)
library(tidyr)


learn_obj_3 <- learn_obj_2 |>
  # 1. Temporarily pivot columns 1 to 13 into a single long column
  pivot_longer(
    cols = starts_with("comp_code"),
    names_to = "Original_Column",
    values_to = "Competency Code"
  ) |>
  mutate("Competency Code" = str_trim(`Competency Code`, "both")) |> 
  # 2. Do the single, clean join
  left_join(
    comp_2, 
    by = "Competency Code"
  ) |>
  # 3. Pivot it back to the original wide format
  pivot_wider(
    names_from = Original_Column,
    values_from = c(`Competency Code`, `Competency Description`),
    names_glue = "{.value}_{Original_Column}"
  ) |> 
  select(-starts_with("Competency Code_"))


```

#### Merge Cleaning

- There are duplicate learning objectives due to varying competencies
- Thus, I set up a group_id column and then bind similar rows

```{r}
#| label: bind-learn-objv
#| message: false
#| warning: false
#| echo: true  

learn_obj_4 <- learn_obj_3 |> 
  group_by(`Learning Objective`) |> 
  summarize(
    across(
      contains("comp_code"), 
      ~ paste(.x, collapse = "; "),
      .names = "{.col}_objs" # e.g. comp_code1_objs, comp_code2_objs
    )
  )

```

#### Bind comp_codes together

```{r}
#| label: bind-comps-together
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

learn_obj_5 <- learn_obj_4 |> 
  unite(
    col = contains()
  )
```
