---
title: "FFAR Project Text Mining - Import Trial"
author: "Patrick Torralba"
format: 
  html:
    theme: flatly
    fontsize: 0.8em
    code-fold: FALSE
    code-tools: TRUE
    code-copy: TRUE
    toc: TRUE
    toc-depth: 4
    smooth-scroll: TRUE
    embed-resources: TRUE
    fig-align: center
editor: visual
---

This import trial is needed bc it will guide how I create the for-loop.

Thus, I will only import 1 .pdf.

# Library
```{r}
#| label: library
#| message: FALSE
#| warning: FALSE
#| echo: TRUE 

library(tidyverse) #clean data
library(pdftools)
```


# Import data
```{r}
#| label: import
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

report_1 <- pdf_text("Old Format Reports/22-000285_Final Report1.pdf")
```

# Clean data

## Data Transformation

Convert from pdf to table

```{r}
#| label: pdf-to-table
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

report_1_v2 <- report_1 |> 
  str_split("\n") |> ## create rows based on the identations, spaces, and tabs
  unlist() |> ## need to do this as necessary step. idk why
  as_tibble()

```

## Indicator Extractor

Due to the formatting of the FFAR impact report, I need to extract information *before* and *after* the general information section

### Pre-General Information Extraction

This section refers to all information occurring before the numbered questions.

```{r}
#| label: pre-gen-info
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  


```

